An Efficient Forecasting Approach\\ to Reduce Boundary Effects in Real-Time Time-Frequency Analysis

### 2020-09-02 Notes/BoundEffRed.tex
code
edit paper

\title{An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects\\ in Time-Frequency Representations}

@@

Attempts to minimize the boundary effects generally consists in softening the discontinuity on signal edges. Usually the approaches can be classified into two main classes, choosing a proper analysis window and extending/forecasting the signal. In the first class, judicious choice of analysis window whose support does not interfere with the boundary points can minimize the occurrence of aberrant patterns near the boundaries of the TF plane, for instance, \cite{Chui92wavelets,Depczynski99fast}. Due to the specific relationship of chosen analysis windows and TF analysis tool, these techniques do not make it possible to reduce the boundary effects of any TF representations. 
%
Another natural idea, the second class, consists of carrying out a preliminary step of extending, or forecasting, the signal beyond its boundaries. Due to its flexibility, various forecasting schemes have been proposed. For example, there exist simple extension schemes that do not take into account the dynamical behavior of the signal, such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet,Chen95symmetric}, or polynomial extrapolation~\cite{Williams97discrete}. 
%
There exist extension schemes based on physically relevant dynamical models, such as the Extended Dynamic Mode Decomposition~\cite{Williams15data} (EDMD) and the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR). There also exist extension schemes based on stochastic models, such as the Trigonometric, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS) algorithm~\cite{DeLivera11forecasting}, the dynamic linear models \cite{west2006bayesian}. 
%
In the physically relevant dynamical models, the oscillation or trend, are usually modeled as the mean of a stochastic process, while in the stochastic models, the oscillation or trend are modeled by the covariance structure of the stochastic process. 
It is also possible to consider polynomial regression \cite{fan1996local} or modeling the mean by splines \cite{hall2005theory}, kernel functions \cite{chang2010training}, and wavelets \cite{marron1998exact}, or nonparametric regression \cite{fan1996local}, to estimate the mean function and hence forecasting the signal. Neural network models, such as the long short-term memory model \cite{vlachas2018data}, is another approach. 
%
The above is a far from exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic. 
%
While the model-based extension scheme gives better-extended signals than those simple extension without dynamical models, they generally have a great computational cost.

@@

Thus, assuming that the quality of the extension step is sufficient, the analysis results obtained that way will be less sensitive to the boundary effects than the result of the analysis tool applied directly to the nonextended signal. We claim and prove that forecasting oscillatory signals based on the simple dynamical model combined with the simple least square approach is sufficient for reducing the boundary effects for the TF analysis, or other kernel-based analysis. See the bottom right of Fig.~\ref{fig:ex.intro} for a snapshot of the result. The main benefit of this simple approach is a numerically efficient solution with a theoretical guarantee for real-time analysis purposes.

>>>>

\title{An Efficient Forecasting Approach\\ to Reduce Boundary Effects in Real-Time Time-Frequency Analysis}

@@

Attempts to minimize the boundary effects generally consists in softening the discontinuity on signal edges. Usually the approaches can be classified into two main classes, choosing a proper analysis window and extending/forecasting the signal. In the first class, judicious choice of analysis window whose support does not interfere with the boundary points can minimize the occurrence of aberrant patterns near the boundaries of the TF plane, for instance, \cite{Chui92wavelets,Depczynski99fast}. Due to the specific relationship of chosen analysis windows and TF analysis tool, these techniques do not make it possible to reduce the boundary effects of all TF representations. 
%
Another natural idea, the second class, consists of carrying out a preliminary step of extending, or forecasting, the signal beyond its boundaries. Due to its flexibility, various forecasting schemes have been proposed. For example, there exist simple extension schemes that do not take into account the dynamical behavior of the signal, such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet,Chen95symmetric}, or polynomial extrapolation~\cite{Williams97discrete}. 
%
There exist extension schemes based on physically relevant dynamical models, such as the Extended Dynamic Mode Decomposition~\cite{Williams15data} (EDMD) and the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR). There also exist extension schemes based on stochastic models, such as the Trigonometric, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS) algorithm~\cite{DeLivera11forecasting}, the dynamic linear models \cite{west2006bayesian}. 
%
In the physically relevant dynamical models, the oscillation or trend, are usually modeled as the mean of a stochastic process, while in the stochastic models, the oscillation or trend are modeled by the covariance structure of the stochastic process. 
It is also possible to consider polynomial regression~\cite{fan1996local} or modeling the mean by splines~\cite{hall2005theory}, kernel functions~\cite{chang2010training}, and wavelets~\cite{marron1998exact}, or nonparametric regression~\cite{fan1996local}, to estimate the mean function before forecasting the signal. Neural network models, such as the long short-term memory model~\cite{vlachas2018data}, is another approach. 
%
The above is a far from exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic. 
%
While the model-based extension scheme gives better-extended signals than those simple extension without dynamical models, they generally have a great computational cost.

@@

Thus, assuming that the quality of the extension step is sufficient, the analysis results obtained that way will be less sensitive to the boundary effects than the result of the analysis tool applied directly to the nonextended signal. We claim and prove that forecasting oscillatory signals based on the simple dynamical model combined with the simple least square approach is sufficient for reducing the boundary effects for the TF analysis, or other kernel-based analysis. To our knowledge, such theoretical analysis does not exist in the literature. See the bottom right of Fig.~\ref{fig:ex.intro} for a snapshot of the result. The main benefit of this simple approach is a numerically efficient solution with a theoretical guarantee for real-time analysis purposes.

### 2020-09-02 Notes/algorithm.tex
edit paper

\paragraph{Dynamical Model and Forecasting} 
Establishing a dynamical model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
\[
\bY = f(\bX)\ .
\]
In a general framework, forecasting means estimating the function $f$ from the observed values taken by the signal, in order to predict its future values. For instance, the dynamic mode decomposition~\cite{Schmid10dynamic,Williams15data} or other more complicated models \cite{Roberts13Gaussian,DeLivera11forecasting,west2006bayesian,vlachas2018data}, allow this by setting additional constraints on the behavior of $f$. We will see, in section~\ref{se:theoretical}, that considering such a complex dynamic model is not necessary for the study of the oscillatory signals of interest to us. That is why we consider here a naive dynamical model, assuming that we have the following relation:
\begin{equation}
\bY = \bA\bX\ ,
\label{eq:dyn.model}
\end{equation}
where $\bA\in\RR^{M\times M}$. In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization of a nonlinear system}, when the system is sufficiently regular. Notice that this linearized dynamical model can be written equivalently in function of the subsignals $\bx_k$, as:
\begin{equation}
\bx_{k+1} = \bA\bx_{k},\ ,\forall k\in\{0,\dots,K-1\}\ .
\end{equation}

The forecasting method consists in estimating the unknown matrix $\bA$. Indeed, let $\tilde\bA$ denotes the estimate of $\bA$. We then obtain the forecasting of the signal at time $\frac{N-1+\ell}{\fs}$ by:
\begin{equation}
\label{eq:prediction}
\tilde \bx [N-1+\ell] = \balpha^{(\ell)}\bx_{K}\ ,
\end{equation}  
where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$; that is to say,
\begin{equation}
\balpha^{(\ell)}=\be_M^T\tilde\bA^\ell\ ,
\label{eq:alpha.l}
\end{equation}
where $\be_M$ is the unit vector of length $M$ given by $\be_M = \begin{pmatrix} 0 & \cdots & 0 & 1\end{pmatrix}^T$.

>>>>

\paragraph{Dynamical Model} 
Establishing a dynamical model consists of determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
\begin{equation}
\bY = f(\bX)\ .
\label{eq:generic.model}
\end{equation}
In a general framework, forecasting means estimating the function $f$ from the observed values taken by the signal, in order to predict its future values. For instance, the dynamic mode decomposition~\cite{Schmid10dynamic,Williams15data} or other more complicated models \cite{Roberts13Gaussian,DeLivera11forecasting,west2006bayesian,vlachas2018data}, allow this by setting additional constraints on the behavior of $f$. We will see, in section~\ref{se:theoretical}, that considering such a complex dynamic model is not necessary for the study of the oscillatory signals of interest to us. That is why we consider here a naive dynamical model, assuming that we have the following relation:
\begin{equation}
\bY = \bA\bX\ ,
\label{eq:dyn.model}
\end{equation}
where $\bA\in\RR^{M\times M}$. In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization of a nonlinear system}, when the system is sufficiently regular. Notice that this linearized dynamical model can be written equivalently in function of the subsignals $\bx_k$, as:
\begin{equation}
\bx_{k+1} = \bA\bx_{k},\ ,\forall k\in\{0,\dots,K-1\}\ .
\end{equation}
From the Takens' embedding perspective~\cite{Takens81detecting}, $\bX$ can be viewed as a lag map. Under the manifold assumption of the intrinsic phase space that hosts the dynamics, $f$ describes the dynamics supported on the manifold space, and $\bA$ approximates the dynamics on the manifold.

\paragraph{Forecasting}
The forecasting method consists in estimating the unknown matrix $\bA$. Indeed, let $\tilde\bA$ denotes the estimate of $\bA$. We then obtain the forecasting of the signal at time $\frac{N-1+\ell}{\fs}$ by:
\begin{equation}
\label{eq:prediction}
\tilde \bx [N-1+\ell] = \balpha^{(\ell)}\bx_{K}\ ,
\end{equation}  
where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$; that is to say,
\begin{equation}
\balpha^{(\ell)}=\be_M^T\tilde\bA^\ell\ ,
\label{eq:alpha.l}
\end{equation}
where $\be_M$ is the unit vector of length $M$ given by $\be_M = \begin{pmatrix} 0 & \cdots & 0 & 1\end{pmatrix}^T$.
